A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-24 paper/ms.tex
update ms (add quartile phased lc)

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler space telescope because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
  using Gaia kinematics, TESS stellar rotation periods, and
  spectroscopic lithium abundances that Kepler\,1627 is a member of
  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
  Kepler\,1627Ab is the youngest planet with a precise age yet found
  by the main Kepler mission.  From unresolved starspot-crossing
  events in the Kepler photometry and a Keck/HIRES transit we find
  that the planet is likely on a prograde orbit.  The average Kepler
  transit profile also appears asymmetric; our current best
  explanation is that of a dusty atmospheric outflow being present,
  but this remains to be confirmed.  Broadly speaking, the
  $\delta$~Lyr cluster is one of many stellar groups whose properties
  have been clarified using Gaia; many other exoplanet hosts belong to
  these new clusters.  Verifying the results of Gaia clustering
  analyses using TESS rotation periods offers a promising path toward
  expanding the census of age-dated planets, and our corresponding
  understanding of planetary evolution.
\end{abstract}

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$~Lyr cluster.} Colored
		points are kinematically selected candidate $\delta$\,Lyr cluster
		members.  Both Kepler (gray panels) and TESS observed portions of
		the cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray;
		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
		entirety of the Kepler mission.  Colors of candidate cluster
		members indicate the duration of TESS observations to date, in
		units of lunar months.
		\label{fig:skychart}
	}
\end{figure}

@@

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:} Color-absolute
    magnitude diagram of candidate \cn\ members, in addition to stars
    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr;
    \citealt{dahm_2015}), and the Gaia EDR3 Catalog of Nearby Stars
    (gray background).  The zoomed right panel highlights the
    pre-main-sequence.  The \cn\ and IC\,2602 are approximately the
    same isochronal age.  {\it Bottom:} TESS and Kepler stellar
    rotation period versus dereddened Gaia color, with the Pleiades
    and Praesepe (650\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most candidate
    \cn\ members are gyrochronally younger than the Pleiades; outliers
    are probably field interlopers.
   \label{fig:age}
	}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

@@

Figures~\ref{fig:phasequarter} and~\ref{fig:phaseyear} show the Kepler
data binned over quarters and years.  In
Figure~\ref{fig:phasequarter}, Quarter 6 shows the strongest asymmetry
out of the quarters: a deviation of about 3\,ppt from expectation.
Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
correspondingly shows the strongest anomaly out of any year in
Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
in each of Years 2, 3, and 4.

We therefore considered four possible explanations for the anomaly:
gravity darkening, transit timing variations, spot-crossing events,
and a persistent asymmetric dusty outflow.

@@

The scenario of transit timing variations (TTVs) producing the
asymmetry seems unlikely because our analysis showed that transit
timing variations in \sn\ have amplitudes smaller than $\approx$10
minutes (see Figure~\ref{fig:obliquity}).  This is consistent with
previous results from \citet{holczer_transit_2016}.  Moreover, the
transit timing variations we do observe are correlated with the local
light curve slope, which on average increases as much as it decreases.
From our analysis, the mean TTV and its standard deviation are
$0.00\pm12.05$ minutes; similarly the mean local slope and its
standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
therefore no expectation for TTVs to produce the asymmetry.

The third possibility is that of starspot crossings.  Young stars 
have higher spot-covering fractions than old stars ({\it
e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
may also host dark starspots at high stellar latitudes \citep[{\it
e.g.}, EK~Dra;][]{strassmeier_starspots_2009}, though interferometric
imaging of spotted giant stars has shown different starspot latitude
distributions than those inferred from Doppler imaging
\citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
spot-crossing anomalies to add coherently over the 144 Kepler
transits, it seems likely that we would need either for spots to be
persistent at a particular latitude (and for the planetary orbit to be
quite misaligned), or for a ``stroboscopic'' longitudinal phasing
\citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
orb}/P_{\rm rot} \approx 2.76$, which means that every 4 transits and
11 stellar rotations, the planet crosses over roughly the same stellar
longitude, which might enable the necessary phasing if the spot-groups
are large and long-lived.  Unfortunately, the  S/N per Kepler transit
is $\approx8$, which renders individual spot-crossing events
unresolved.  This explanation seems marginally plausible, because the
expected spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$)
resemble the observed amplitude of the asymmetry ($\approx 50\,{\rm
ppm}$).  We ultimately disfavor it though, because {\it a)} it seems
implausible that the starspot crossings could add coherently, and {\it
b)} there is no reason to expect starspot crossing events to last
exactly half the transit
duration.

>>>>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler space telescope because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
  using Gaia kinematics, TESS stellar rotation periods, and
  spectroscopic lithium abundances that Kepler\,1627 is a member of
  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
  Kepler\,1627Ab is the youngest planet with a precise age yet found
  by the main Kepler mission.  The Kepler data show two notable
  anomalies.  First, after removal of the stellar variability, the
  average Kepler transit profile appears to be asymmetric.  One
  possible explanation is that a dusty atmospheric outflow is present,
  but this remains to be confirmed.  Second, the individual transits
  show a correlation between the transit time and the local slope of
  the light curve.  Interpreted in a framework in which unresolved
  starspot-crossing events cause the transit timing variations, this
  suggests that the planet is on a prograde orbit.  We present
  Keck/HIRES observations that are consistent with the latter point, though
  the data do not have the precision necessary to claim a detection of the
  spectroscopic transit.  More broadly, the $\delta$~Lyr cluster
  is one of many stellar groups whose properties have been clarified
  using Gaia; many other exoplanet hosts belong to these new clusters.
  Verifying the results of Gaia clustering analyses using TESS
  rotation periods offers a promising path toward expanding the census
  of age-dated planets, and our corresponding understanding of
  planetary evolution.
\end{abstract}

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$~Lyr cluster.} Colored
		points are kinematically selected candidate $\delta$\,Lyr cluster
		members.  Both Kepler (gray panels) and TESS observed portions of
		the cluster.  Stars visible by eye ($m_{\rm V}<6.5$) are shown in gray;
		Deneb, Vega, and $\delta$~Lyr are orange.  Kepler\,1627 (green star) was observed during the
		entirety of the Kepler mission, and to date has been observed for
		two lunar months by TESS.%.  Colors of candidate cluster
		%members indicate the duration of TESS observations to date, in
		%units of lunar months.
		\label{fig:skychart}
	}
\end{figure*}

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.7\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:} Color-absolute
    magnitude diagram of candidate \cn\ members, in addition to stars
    in IC\,2602 ($\approx38$\,Myr), the Pleiades ($\approx 115$\,Myr;
    \citealt{dahm_2015}), and the Gaia EDR3 Catalog of Nearby Stars
    (gray background).  The zoomed right panel highlights the
    pre-main-sequence.  The \cn\ and IC\,2602 are approximately the
    same isochronal age.  {\it Bottom:} TESS and Kepler stellar
    rotation period versus dereddened Gaia color, with the Pleiades
    and Praesepe (650\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.  Most candidate
    \cn\ members are gyrochronally younger than the Pleiades; outliers
    are probably field interlopers.
   \label{fig:age}
	}
\end{figure*}

@@

		\vspace{-0.2cm}	
		\subfloat{
			\includegraphics[width=0.5\textwidth]{f5a.pdf}
			\includegraphics[width=0.5\textwidth]{f5b.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
    The Kepler data span 1{,}437 days (3.9 years), sampled at
    30 minute cadence;  a 100 day segment is shown.  The
    top panel shows the \texttt{PDCSAP} median-subtracted flux in
    units of parts-per-thousand ($\times 10^{-3}$).  The dominant
    signal is induced by starspots.  The stellar
    variability model (orange line) is subtracted below, revealing the
    transits of \pn.  The
    online Figure Set spans the entire 3.9 years of observations.
    {\it Bottom}:
    Phase-folded transit of Kepler 1627Ab with stellar variability
    removed.  Windows over 20 hours ({\it left}) and the entire orbit
    ({\it right}) are shown, and the residual after subtracting the
    transit in the bottom-most row.  The 2-$\sigma$ model uncertainties
    and the best-fit model are the light purple band and the dark
    purple line.  Gray points are individual flux measurements; black
    points bin these to 20 minute intervals, and have a representative
    1-$\sigma$ error bar in the lower right of each panel.  The
    asymmetric residual during transit appears to be robust to
    different detrending methods.
    \label{fig:lc}
  }
\end{figure*}

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=\textwidth]{f19.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Transit models and residuals, binned by local slope of the
    raw light curve}.  Local linear slopes are measured over each
    transit window (144 transits total).  Four outlier transits were
    removed, leaving 140 transits.  These were then divided into
    quartiles, so that each panel shows 35 transits binned together.
    The exact light curve slopes are listed in the lower left panels.
    Representative uncertainties for the black points (binned at 20
    minute intervals) are shown in the lower right of each panel.  A
    similar transit asymmetry to that shown in Figure~\ref{fig:lc}
    seems to be present in three of the four bins.
		\label{fig:phaseslope}
	}
\end{figure}

@@

As a means of exploring the robustness of the asymmetry,
Figures~\ref{fig:phasequarter},~\ref{fig:phaseyear},
and~\ref{fig:phaseslope} show the Kepler data binned in three ways:
over quarters, years, and local slope quartiles.  Over quarters
(Figure~\ref{fig:phasequarter}), Quarter 6 shows the strongest asymmetry
out of any quarter: a deviation of about 3\,ppt from expectation.
Quarter 7 shows an anomaly at roughly the same transit phase.  Year 2
correspondingly shows the strongest anomaly out of any year in
Figure~\ref{fig:phaseyear}; the asymmetry is visually apparent however
in each of Years 2, 3, and 4.
Binned by local slope quartiles (Figure~\ref{fig:phaseslope}), the
asymmetry is visually present in three of the four quartiles: the only
bin for which it appears to not be present is ${\rm d}f/{\rm
d}t\in[0.001,0.033]\,{\rm ppt\,day}^{-1}$.

We considered four possible interpretations of the transit asymmetry:
gravity darkening, transit timing variations, spot-crossing events,
and a persistent asymmetric dusty outflow.

@@

The scenario of transit timing variations (TTVs) producing the
asymmetry seems unlikely because our analysis showed that transit
timing variations in \sn\ have amplitudes smaller than $\approx$10
minutes (see Figure~\ref{fig:obliquity}).  This is consistent with
previous results from \citet{holczer_transit_2016}.  Moreover, the
transit timing variations we do observe are correlated with the local
light curve slope, which on average increases as much as it decreases.
From our analysis, the mean TTV and its standard deviation are
$0.00\pm12.05$ minutes; similarly the mean local slope and its
standard deviation are $0.49\pm44.49$\,ppt\,day$^{-1}$.  There is
therefore no expectation for TTVs to produce the asymmetry.  A
separate line of argument comes from Figure~\ref{fig:phaseslope}.  If
the local slope were essential to producing the transit asymmetry, we
would expect that in the largest ${\rm d}f/{\rm d}t$ bin, ${\rm
d}f/{\rm d}t\in[0.033,0.092]\,{\rm ppt\,day}^{-1}$, the sign of the
asymmetry would reverse.  We do not see evidence for that being the
case.

The third and related possibility is that of starspot crossings.
Young stars have higher spot-covering fractions than old stars ({\it
e.g.}, \citealt{morris_relationship_2020}).  Young solar-type stars
may also host dark starspots at high stellar latitudes \citep[{\it
e.g.}, EK~Dra;][]{strassmeier_starspots_2009}, though interferometric
imaging of spotted giant stars has shown different starspot latitude
distributions than those inferred from Doppler imaging
\citep{roettenbacher_contemporaneous_2017}.  Regardless, for any
spot-crossing anomalies to add coherently over the 144 Kepler
transits, it seems likely that we would need either for spots to be
persistent at a particular latitude (and for the planetary orbit to be
quite misaligned), or for a ``stroboscopic'' longitudinal phasing
\citep[{\it e.g.},][]{dai_stellar_2018}.  For our system, $P_{\rm
orb}/P_{\rm rot} \approx 2.76$, which means that every 4 transits and
11 stellar rotations, the planet crosses over roughly the same stellar
longitude, which might enable the necessary phasing if the spot-groups
are large and long-lived.  Unfortunately, the  S/N per Kepler transit
is $\approx8$, which renders individual spot-crossing events
unresolved.  This explanation seems marginally plausible, because the
expected spot-crossing anomaly amplitudes ($\approx 100\,{\rm ppm}$)
resemble the observed amplitude of the asymmetry ($\approx 50\,{\rm
ppm}$).  We ultimately disfavor it though, because {\it a)} it seems
implausible that the starspot crossings could add coherently, and {\it
b)} there is no reason to expect starspot crossing events to last
exactly half the transit duration.
